---
title: "Aggregate Loss Models"
author: "A short course authored by the Actuarial Community"
date: "21 Jan 2021"
output:
  beamer_presentation:
    includes:
      in_header: preamble.tex
  ioslides_presentation: default
  slidy_presentation: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



#  Introduction

## Basic Terminology

-   \textcolor{blue}{Loss} - amount of financial losses suffered by the insured
-   \textcolor{blue}{Claim} - indemnification upon the occurrence of an insured event. Some authors use claim and loss interchangeably 
-   \textcolor{blue}{Frequency} - how often an insured event occurs (claim count) in one period (typically six months or one year) 
-   \textcolor{blue}{Severity} - Amount, or size, of losses for an insured event 
-   \textcolor{blue}{Aggregate Loss} - Total amount paid for a defined set of insureds in one period (typically six months or one year)
 
 
## Goal
 
-  Build a model for the total payments by an insurance system in a fixed time period
-  The insurance system could be a single policy or a portfolio of policies
-  Frequency and severity models are building blocks


## Models
 
-  Two ways to build a model for aggregate losses on a defined set of insurance contracts
-  Individual Risk Model: record losses for each contract and then add them up
-  Collective Risk Model (a.k.a. compound model): record losses as claims are made and then add them up


## Example
 
-  An insurance portfolio of three policies:
\begin{center}
\begin{tabular}{cccc}\hline
    Policy ID & Claim ID & Loss Amount \\ \hline
      1 & - & - \\
      2 & 1 & 10 \\
      3 & 1 & 10 \\
      3 & 2 & 10 \\  
      4 & 1 & 10 \\
      4 & 2 & 10 \\  
      4 & 3 & 10 \\  \hline
  \end{tabular}
\end{center}

\bigskip
 
-  Aggregate losses:  
   -  Individual Risk Model: $0 + 10 + 20 +30 = 60$
   -  Collective Risk Model: $10 + 10 + 10 + 10 + 10 + 10= 60$


## Applications
 
-  Actuarial applications of aggregate loss models
   -   Ratemaking
   -   Capital management
   -   Risk financing


## Summary
 
In this section, we learned how to:

- Record aggregate losses from an insurance system
- Identify actuarial applications of aggregate loss models 

   
   

# Individual Risk Model

## Individual Risk Model
 
The **individual risk model** represents the aggregate loss as a sum of a fixed number of insurance contracts
$$
S =X_1 + \ldots + X_n,
$$
where
  
-    $S$ denotes the aggregate loss for $n$ (a fixed number) contracts
-    $X_i$ denotes the loss for the $i$th contract for $i=1,\ldots,n$
-    $X_i$ are assumed to independent but are not necessarily identically distributed, due to different coverage or exposure
-    $X_i$ usually has a probability mass at zero
  
  

## Applications
 
-  Originally developed for life insurance
   -  Probability of death within a year is $q_i$;
   -  Fixed benefit paid for the death of the $i$th person is $b_i$.
-   The distribution of the loss to the insurer for the $i$th policy is

$$
\begin{array}{ll}
f_{X_i}(x) = \left\{\begin{array}{ll}
                               1-q_i, & x=0 \\
                               q_i, & x=b_i
                             \end{array}
             \right.
\end{array}
$$


## Applications
 
-   Two-part framework
$$
X_i = I_i\times B_i = \left\{\begin{array}{ll}
                               0, & I_i=0 \\
                               B_i, & I_i=1
                             \end{array} \right.
$$

-   $I_1,\ldots,I_n$, $B_1,\ldots,B_n$ are independent.
-   $I_i$ is an indicator (Bernoulli) that is 1 with probability $q_i$ and 0 with probability 1-$q_i$. It indicates whether the $i$th policy has a claim.
-   $B_i$, a r.v. with nonnegative support, represents the amount of losses of policy $i$, given that a claim is made. It can follow a degenerate distribution such as the life insurance example.




## Example
 
Consider the example of Wisconsin Property Fund. 

```{r}
Insample <- read.csv("Insample.csv", header = T)
Insample2010 <- subset(Insample, Year==2010)

I <- 1*(Insample2010$Freq>0)
B <- Insample2010[which(Insample2010$Freq>0),]$y 
```

```{r}
length(I)
length(B)
```


## Example
 

```{r}
table(I)

summary(B)
```



## Aggregate Loss Distribution
 
The distribution of aggregate loss $S$ is defined as:
$$F_S(s) = {\rm Pr}(X_1 + X_2 + \cdots +X_n \le s)$$

-   In general, it is  difficult to evaluate
-   When $X_i$, $i=1,\ldots,n$, are i.i.d., it is known as $n$-fold convolution of cdf of $X$ 
 

## R Example #1


```{r,warning = FALSE,message = FALSE,out.width = '0.6\\maxwidth',fig.align='center'} 
library(distr)
X <- Exp(rate=1); Y <- Lnorm(meanlog=1,sdlog=1)
conv <- convpow(X+Y,1)             
f.Z  <- d(conv)
z <- seq(0,30,0.01)
plot(z,f.Z(z), col="blue",type="l")
``` 


## R Example #2


```{r,warning = FALSE,message = FALSE,out.width = '0.6\\maxwidth',fig.align='center'} 
library(distr)
X <- Exp(rate=1)
conv <- convpow(X,10)             
f.Z  <- d(conv)
z <- seq(0,30,0.01)
plot(z,f.Z(z), col="blue",type="l")
``` 





## Summary

In this section, we learned how to:

- Build an individual risk model for a portfolio of insurance contracts
- Apply individual risk model to life and nonlife insurance 
- Compute the distribution of aggregate losses from an individual risk model


 




# Collective Risk Model - Part I

## Collective Risk Model
 
   The **collective risk model** has representation
$$
S = X_1 + \ldots + X_N,
$$
with $S$ being the aggregate loss of $N$ (a random number) individual claims $(X_1,\ldots,X_N)$.


-  Key assumptions
   -    Conditional on $N=n$, $X_1,\ldots,X_n$ are i.i.d. random variables
   -    The distribution of $N$ and the common distribution of $X$ are independent of each other.
- Two building blocks: frequency $N$ and severity $X$

 
## Compound Distribution
 
The cumulative distribution function (cdf) is $F_S(s)$. The probability density function (pdf) or probability mass function (pmf) is $f_S(s)$. Specifically,
   
$$
\begin{array}{ll}
   F_S(s) &= {\rm Pr}(S\leq s)\\
   &= \sum_{n=0}^{\infty}{\rm Pr}(N=n)\cdot {\rm Pr}(S\leq s|N=n)\\
   &=\sum_{n=0}^{\infty}{\rm Pr}(N=n)\cdot F_X^{*n}(s)\\ \\
   f_S(s)&=\sum_{n=0}^{\infty}{\rm Pr}(N=n)\cdot f_X^{*n}(s)
   \end{array}
$$
 
## Moments
 
$$
S = X_1 + \ldots + X_N
$$

Using the law of iterated expectations to calculate the mean
$$
  {\rm E}(S) = {\rm E}[{\rm E}(S|N)] = {\rm E}[N{\rm E}(X)] = {\rm E}(N){\rm E}(X)
$$

Using the law of total variation to calculate the variance
$$  
\begin{array}{ll}
  {\rm Var}(S) &= {\rm E}[{\rm Var}(S|N)]+ {\rm Var}[{\rm E}(S|N)] \\
  &= {\rm E}[N{\rm Var}(X)] + {\rm Var}[N{\rm E}(X)]\\
  &={\rm E}(N){\rm Var}(X) + {\rm Var}(N){\rm E}(X)^2
  \end{array}
$$

---
## Moments - Example

Consider model $S = X_1 + \ldots + X_N$ where

- $N \sim Poisson(\lambda)$
- $X \sim Gamma(shape = 1/\phi, scale=\phi\mu)$

Using facts ${\rm E}(N) = {\rm Var}(N) = \lambda$, ${\rm E}(X)=\mu$, and ${\rm Var}(X)=\phi\mu^2$, one has
$$
  {\rm E}(S) = {\rm E}(N){\rm E}(X) = \lambda\mu
$$
and
$$  
\begin{array}{ll}
  {\rm Var}(S) &={\rm E}(N){\rm Var}(X) + {\rm Var}(N){\rm E}(X)^2 = \lambda\mu^2(1+\phi)
  \end{array}
$$
---

 
## Model Fitting
 
The assumptions suggest that we can build an aggregate loss model, the compound model, in three steps: 
  
1.    Develop a model for the frequency distribution of $N$, the primary distribution, based on data 
2.    Develop a model for the severity distribution of $X$, the secondary distribution, based on data 
3.    Using these two models, carry out the necessary calculations to obtain the distribution of $S$
 

## Model Fitting

Consider the example of Wisconsin Property Fund. 


```{r}
Insample <- read.csv("Insample.csv", header = T)
Insample2010 <- subset(Insample, Year==2010)
```

Number of policyholders: 
```{r}
nrow(Insample2010)
```





## Model Fitting

The frequency model: $N$


```{r,R.options = list(width = 50)}
N <- Insample2010$Freq
```


```{r,R.options = list(width = 50)}
table(N)
```


  
## Model Fitting



The severity model: $\bar{X}=(X_1+\ldots+X_n)$ for $N=n>0$

```{r,R.options = list(width = 50)}
Xbar <- Insample2010$yAvg[which(Insample2010$Freq>0)]
```


```{r,R.options = list(width = 50)}
summary(Xbar)
```

  
  
  
## Summary

In this section, we learned how to:

- Build a collective risk model for a portfolio of insurance contracts
- Calculate mean and variance of the aggregate loss 
- Fit frequency and severity components in a collective risk model





# Collective Risk Model - Part II



## Computing the Aggregate Loss Distribution
 
- Consider a collective risk model $$S=X_1+\cdots+X_N$$

- Several numerical strategies:
  
  -   **direct calculation**: difficult part is the evaluation of $n$-fold convolutions
  -   **recursive method**: considerable savings in computation time
  -   **Monte Carlo simulation**: approximation using the empirical distribution
  

- `R` implementation: package `actuar`



## Direct Calculation
 
- The distribution of $S$ can be calculated using:
\begin{align*}
F_S(s) & ={\Pr(S\le s)} \\
& =\sum_{n=0}^{\infty}{\rm Pr}(N=n)\cdot {\rm Pr}(S \le s|N=n)\\
& =\sum_{n=0}^{\infty}{\rm Pr}(N=n)\cdot F_X^{*n}(s)
\end{align*}


- To compute the distribution in `R`, one has to discretize the severity distribution so that it has support $\left\{0,1,\ldots,m\right\}$ 



## Direct Calculation
 
The frequency and severity distributions are summarized by:
$$
\begin{array}{lccccc}
     \hline\hline
     N & 0 & 1 & 2 & 3 & 4  \\
     f_N(n) & 0.2 & 0.2& 0.2& 0.2 & 0.2 \\
     \hline
     X & 50 & 100 & 150 & 250  & \\
     f_X(x) & 0.2 & 0.3 & 0.4 & 0.1 & \\
     \hline\hline
\end{array}
$$
The discretized severity distribution is 
$$
\begin{array}{lcccccc}
     \hline\hline
     X & 0 & 1  & 2 & 3  &  4 & 5\\
     f_X(x) & 0 & 0.2 & 0.3 & 0.4 & 0& 0.1 \\
     \hline\hline
\end{array}
$$


## Direct Calculation
 
```{r, message=FALSE, warning=FALSE}
library(actuar)
pn <- rep(0.2,5)
fx <- c(0,0.2,0.3,0.4,0,0.1)
Fs <- aggregateDist(method="convolution", 
                    model.freq = pn,
                    model.sev = fx, x.scale = 50)
summary(Fs)
```

## Direct Calculation
 
```{r, message=FALSE, warning=FALSE, out.width= '0.9\\maxwidth',fig.align='center'}
plot(Fs)
```


## Recursive Method

- Recursive method assumes:
  - Frequency $N$ satisfies $p_n = (a+b/n)p_{n-1}$ for $n=\textcolor{blue}{1},2,3,\ldots$
  - Severity $X$ has a distribution distribution $f_X$ on support $\left\{0,1,\ldots,m\right\}$

\bigskip

- The probability function of $S$ can be calculated using:
\begin{align*}
 f_S(s) &= {\rm Pr}(S=s)\\ 
&= \frac{1}{1-af_X(0)}\sum_{x=1}^{s\wedge m}\left(a+\frac{bx}{s}\right)f_X(x)f_S(s-x)
\end{align*}

- The method extends to the more general case where frequency $N$ satisfies $p_n = (a+b/n)p_{n-1}$ for $n=2,3,\ldots$

---
\scalefont{0.8}
\begin{align*}
 f_S(s) &= {\rm Pr}(S=s)\\ 
&= \frac{1}{1-af_X(0)}\left\{(p_1-(a+b)p_0)f_X(s) + \sum_{x=1}^{s\wedge m}\left(a+\frac{bx}{s}\right)f_X(x)f_S(s-x)\right\}
\end{align*}
---



## Recursive Method

Consider the example: 

- Frequency $N$ is Poisson with mean 5. 

- Severity
$$
\begin{array}{lcccc}
     \hline\hline
     X & 50 & 100 & 150 & 250  \\
     f_X(x) & 0.2 & 0.3 & 0.4 & 0.1  \\
     \hline\hline
\end{array}
$$


## Recursive Method

```{r, message=FALSE, warning=FALSE,out.width= '0.7\\maxwidth',fig.align='center'}
library(actuar)
fx <- c(0,0.2,0.3,0.4,0,0.1)
Fs <- aggregateDist(method="recursive", 
                    model.freq = "poisson", lambda = 5,
                    model.sev = fx, x.scale = 50)
plot(Fs)
```



## Monte Carlo Simulation

- For $j=1,\ldots, m$, do
  - Generate the number of claims $n_j$ from the frequency distribution $N$
  - Generate $n_j$ claim amount from severity distribution $X$, denoted by $x_{j,1},\ldots,x_{j,n_j}$
  - Calculate aggregate loss $s_j = x_{j,1}+\cdots+x_{j,n_j}$

\bigskip
- The empirical distribution can be calculated as
$$\hat{F}_s(s) = \frac{1}{m} \sum_{i=1}^m I(s_i \le s)$$


## Monte Carlo Simulation

```{r, message=FALSE, warning=FALSE,out.width= '0.7\\maxwidth',fig.align='center'}
library(actuar)
freq <- expression(data = rpois(lambda=3))
sev <- expression(data = rgamma(shape=100, scale=5))
Fs <- aggregateDist("simulation", nb.simul = 2500,
                    model.freq = freq, model.sev = sev)
plot(Fs)
```



## Summary

In this section, we learned how to:

- Compute aggregate loss distribution
- Implement numerical strategies in `R`




 
# Tweedie Distribution

## Tweedie Distribution
 
The Tweedie distribution is defined as a Poisson sum of gamma random variables

$$
    S = (X_1+\cdots+X_N)
$$

where $N\sim Poisson(\omega\lambda)$ and $X\sim Gamma(\alpha,\theta)$.
 
 
 
## Tweedie Distribution with Exposure
 
With exposure, the Tweedie variable is the aggregate losses per unit of expoure:

$$
    S = (X_1+\cdots+X_N)/\textcolor{blue}{\omega}
$$
where 

  - $\omega$ is the exposure
  - $N\sim Poisson(\textcolor{blue}{\omega}\lambda)$ 
  - $X\sim Gamma(\alpha,\theta)$
 
 
  
## Distribution Function for Tweedie Distribution
 
The cdf of $S$ is

$$
\begin{array}{ll}
    F_S(s) &= \sum_{n=0}^{\infty}{\rm Pr}(N=n)\cdot {\rm Pr}(S\leq s|N=n)\\
    & = {\rm Pr}(N=0) + \sum_{n=1}^{\infty}{\rm Pr}(N=n)\cdot {\rm Pr}(S\leq s|N=n)\\
    & = e^{-\omega\lambda} + \sum_{n=1}^{\infty} e^{-\omega\lambda}\frac{(\omega\lambda)^n}{n!}\Gamma\left(n\alpha;\frac{s}{\theta/\omega}\right)
  \end{array}
$$

Note that

$$
S|(N=n)=(X_1+\cdots+X_n)/\omega\sim Gamma(n\alpha,\theta/\omega)
$$
 
## Tweedie Distribution
 
Consider reparameterizations

$$
\lambda=\frac{\mu^{2-p}}{\phi(2-p)},~~\alpha=\frac{2-p}{p-1},~~\theta=\phi(p-1)\mu^{p-1}
$$

   For $p\in(1,2)$, the Tweedie distribution can be presented as:

$$
\begin{array}{ll}
    f_S(s) = \exp\left\{\frac{\omega}{\phi}\left(\frac{-s}{(p-1)\mu^{p-1}}-\frac{\mu^{2-p}}{2-p}\right)+c(s;p,\phi/\omega)\right\}
  \end{array}
$$

and

$$
{\rm E}(S)=\mu,~~{\rm Var}(S)=\frac{\phi}{\omega}\mu^p
$$

## Example

The histogram of annual medical expenses from a randomly selected 5,000 individuals:

```{r,echo=F}
# data generating
lambda = 1
alpha = 4
theta = 500

set.seed(123)
size = 5000
S <- rep(NA,size)
N <- rpois(size,lambda)
for (i in 1:size){
  Y <- rgamma(N[i],shape = alpha, scale = theta)
  S[i] <- sum(Y)
}
dat <- data.frame(ID=1:size,expense=S)
#write.csv(dat, file="MedicalExp.csv",row.names = F)
#xi <- (alpha+2)/(alpha+1)
expense <- dat$expense
hist(expense, main="", prob=T, breaks=50,xlab="Medical Expenses")
```


## Example

```{r,message=FALSE, warning=FALSE}
library(statmod)
p <- 1.2
fit <- glm(expense~1,family=tweedie(var.power=p,
                                    link.power=0))
```

```{r}
mu <- exp(summary(fit)$coefficient[1])
phi <- summary(fit)$dispersion
```


The estimates of $\mu$ is `r round(mu,2)`:

The estimates of $\phi$ is `r round(phi,2)`:

## Example

```{r, eval=FALSE}
library(tweedie)
pct <- seq(0.01,0.99,0.01)
plot(qtweedie(pct,p,mu,phi),quantile(expense,probs=pct),
        xlab="Theoretical Quantile",
        ylab="Empirical Quantile")
abline(0,1)
```


## Example

```{r,echo=F, fig.align="center"}
library(tweedie)
pct <- seq(0.01,0.99,0.01)
plot(qtweedie(pct,p,mu,phi),quantile(expense,probs=pct),
     xlab="Theoretical Quantile",
     ylab="Empirical Quantile",
)
abline(0,1)
```

 
## Summary

In this section, we learned how to:

- Construct the Tweedie distribution from a collective risk model
- Establish the Tweedie distribution as a member of the exponential family of distributions
- Fit Tweedie distribution as a generalized linear model

 

# Effects of Coverage Modifications


## Aggregate Deductible


 
**Definition** Insurance on the aggregate losses, subject to a deductible, is called \textit{stop-loss insurance}. 

-   Assume aggregate loss $S$ and deductible $d$ 
-   The cost of the insurer is 

$$
\begin{array}{ll}
(S-d)_{+} =  \left\{\begin{array}{ll}
                               0, & S \le d \\
                               S-d, & S>d
                             \end{array}
             \right.
\end{array}
$$

 
## Aggregate Deductible


The expected cost of this insurance is called the \textit{net stop-loss premium}.

It can be computed as:


$$
E[(S-d)_{+}] =E(S)-E(S\wedge d) 
$$
where 
 
$$
 {\rm E}(S\wedge d)=\int_0^d [1-F_S(s)] ds 
$$


## Per-occurrence Deductible
 
- Consider aggregate loss $$S=X_1+\cdots + X_N$$

- For a per-occurrence deductible $d$, we examine its effect on 
   -  Claim frequency $N$
   -  Claim Severity $X$





## Per-occurrence Deductible and Frequency

- Let $N$ be the number of losses and $N^{P}$ be the number of payments
- Represent $N^{P}$ as a compound frequency distribution

$$N^{P}= I_1+\cdots+I_{N}$$
where, for $i=1,\ldots,N$,
$$
I_i = \left\{
      \begin{array}{cc}
        1 & {\rm ~with ~probability~} {\rm Pr}(X_i>d)\\
        0 & {\rm ~with ~probability~} {\rm Pr}(X_i \le d)
      \end{array}
    \right.
$$ 
   





## Per-occurrence Deductible and Frequency
 
- The distribution of $N^P$ can be derived using the probability generating function


- Let $v={\rm Pr}(X>d)$, the pgf of $N^{P}$ is

$$
\begin{array}{ll}
    P_{N^{P}}(z)&=P_{N}[P_{I}(z)]\\
    &=P_{N}[1+v(z-1)]
    \end{array}
$$ 




## Per-occurrence Deductible and Frequency
 
-  **Example:** $N\sim Poisson(\lambda)$
  
-  The pgf of $N^{P}$ is

$$
\begin{array}{ll}
   P_{N^{P}}(z)&=\exp\left\{\lambda[\textcolor{blue}{1+v(z-1)}-1]\right\}\\
   &= \exp\left\{\textcolor{blue}{v\lambda}(z-1)\right\}
   \end{array}
$$

-  $N^{P}$ is from the same family, and $N^{P}\sim Poisson(\lambda^{*}=v\lambda)$
 
-  Similar results for other frequency distributions
 
 
 
 
## Per-occurrence Deductible and Severity
 

- Let $X^{P}$ be the amount of payments

- It is defined as

$$
X^P = X - d | X>d
$$ 

- Its distribution is:

$$F_{X^P}(x) = \frac{F_X(x+d)-F_X(x)}{1-F_X(x)}$$

## Per-occurrence Deductible
 
- Ground up loss $$S=X_1+\cdots + X_N$$

- With deductible $d$, the aggregate loss to the insurer is 
$$S=X^P_1+\cdots + X^P_{N^P}$$
  - $N^P$ is the number of payments
  - $X^P$ is the amount of payments
  
  
## Summary

In this section, we learned how to:

- Examine the impact of aggregate deductible on the aggregate loss
- Examine the effect of per-occurrence deductible on frequency and severity components in the aggregate loss

 